---
title: "Introduction to ArchaeoChron"
author: "A. Philippe and M.-A. Vibet"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
    toc: true
vignette: >
  %\VignetteIndexEntry{Introduction to ArchaeoChron}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```

```{r packages}
library(ArchaeoChron)
library(ArchaeoPhases) 
library(coda)
```

# Bayesian model for combining Gaussian dates

## Example of a sunspot  

The sunspot data is constituted of several dates assumed to be contemporaneous of a single event. These dates do not need any calibration but their unit is in year before 2016. 

## Simple model for combining Gaussian dates

Let's first use a simple Bayesian model. 

```{r}
data(sunspot)

MCMC1 <- combination_Gauss(
  M = sunspot$Age[1:5], 
  s = sunspot$Error[1:5], 
  refYear = rep(2016,5), 
  studyPeriodMin = 900, 
  studyPeriodMax = 1500, 
  variable.names = c('theta')
)
```

The output of the function `combination_Gauss()` is a Markov chain of the posterior distribution of the Bayesian model. 

### Convergence of the Markov chain

First, let's check the convergence of the Markov chain of each parameter.

```{r}
plot(MCMC1)
gelman.diag(MCMC1)
```

The Gelman diagnotic gives point estimates about 1, so convergence is reached. 

### Statistics of the posterior distribution of the event.

Now, we can use the package **ArchaeoPhases** in order to describe the posterior distribution of the parameters. 

Here we will focus on the posterior distribution of the date of interest i.e. the parameter 'theta'. 

```{r}
MCMCSample1 <- rbind(MCMC1[[1]], MCMC1[[2]])

MarginalStatistics(MCMCSample1[,1], level = 0.95)
MarginalPlot(MCMCSample1[,1], level = 0.95)
```

The date (mean posterior distribution) of the sunspot estimated using a simple Bayesian model that combines dates, is dates at `r M1[1,1] ` years after Christ. This date is associated with a 95% confidence interval : [`r M1[8,1] `, `r M1[9,1] `].

## Bayesian model for combining Gaussian dates and handling potential outliers

Let's use the function `combinationWithOutliers_Gauss()`.

```{r}
MCMC2 <- combinationWithOutliers_Gauss(
  M = sunspot$Age[1:5],
  s = sunspot$Error[1:5],
  refYear = rep(2016,5),
  outliersIndivVariance = rep(1,5),
  outliersBernouilliProba = rep(0.2, 5),
  studyPeriodMin = 800,
  studyPeriodMax = 1500,
  variable.names = c('theta')
)
plot(MCMC2)
```

The output of the function `combinationWithOutliers_Gauss()` is a Markov chain of the posterior distribution of the Bayesian model. 

Here we will focus on the posterior distribution of the date of interest i.e. the parameter 'theta'. 

```{r}
MCMCSample2 <- rbind(MCMC2[[1]], MCMC2[[2]])
MarginalStatistics(MCMCSample2[,1], level = 0.95)
MarginalPlot(MCMCSample2[,1], level = 0.95)
```

The date (mean posterior distribution) of the sunspot, estimated using a simple Bayesian model that combines dates and allows for outliers, is dates at `r M2[1,1] ` years after Christ. This date is associated with a 95% confidence interval : [`r M2[8,1] `, `r M2[9,1] `].


## Bayesian model for combining Gaussian dates with a random effect

Let's now use the function `combinationWithRandomEffect_Gauss()`.

```{r}
MCMC3 <- combinationWithRandomEffect_Gauss(
  M = sunspot$Age[1:5],
  s = sunspot$Error[1:5],
  refYear = rep(2016,5),
  studyPeriodMin = 0,
  studyPeriodMax = 1500,
  variable.names = c('theta')
)
plot(MCMC3)
```

The output of the function `combinationWithRandomEffect_Gauss()` is a Markov chain of the posterior distribution of the Bayesian model. 

Here we will focus on the posterior distribution of the date of interest i.e. the parameter 'theta'. 

```{r}
MCMCSample3 <- rbind(MCMC3[[1]], MCMC3[[2]])
MarginalStatistics(MCMCSample3[,1], level = 0.95)
MarginalPlot(MCMCSample3[,1], level = 0.95)
```

The date (mean posterior distribution) of the sunspot, estimated using a simple Bayesian model that combines dates and allows for random effects, is dates at `r M3[1,1] ` years after Christ. This date is associated with a 95% confidence interval : [`r M3[8,1] `, `r M3[9,1] `].

## Event Model : Bayesian model for combining Gaussian dates with individual random effects

If we want to date that event, we can use the Event model for combining Gaussian dates. In that example, we will investigate the posterior distribution of the date of the event (called 'theta') and the posterior distribution of the dates associated with this event.

### Generating the Markov chain of the posterior distribution

Finally, let's use the function `eventModel_Gauss()` and the first 10 dates of the dataset sunspot. The study period should be given in calendar year (BC/AD). 

```{r}
MCMC4 <- eventModel_Gauss(
  M = sunspot$Age[1:5],
  s = sunspot$Error[1:5],
  refYear = rep(2016,5),
  studyPeriodMin = 900,
  studyPeriodMax = 1500,
  variable.names = c('theta', 'mu')
)
```

The output of the `eventModel_Gauss()` is the Markov chain of the posterior distribution. 

### Convergence of the Markov chain

First, let's check the convergence of the Markov chain of each parameter.

```{r}
plot(MCMC4)
gelman.diag(MCMC4)
```

The Gelman diagnotic gives point estimates about 1, so convergence is reached. 

### Statistics of the posterior distribution of the event.

Here we will focus on the posterior distribution of the event of interest (parameter 'theta'). 

```{r}
MCMCSample4 <- rbind(MCMC4[[1]], MCMC4[[2]])
MarginalStatistics(MCMCSample4[,6], level = 0.95)
MarginalPlot(MCMCSample4[,6], level = 0.95)
```

The date (mean posterior distribution) of the sunspot, estimated using a simple Bayesian model that combines dates and allows for individual random effects, is dates at `r M4[1,1] ` years after Christ. This date is associated with a 95% confidence interval : [`r M4[8,1] `, `r M4[9,1] `].
